---
title: "R basics"
format: revealjs
editor: visual
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
# Set up the environment
opts_chunk$set(echo=TRUE, eval=FALSE, dpi=100, warnings=FALSE, message=FALSE, warning=FALSE, fig.width = 10, fig.height = 6) 
```

## A simple calculator

```{r}
1 + 1
```

## 'Vectors' as building blocks

```{r}
c(1, 2, 3)
c("January", "February", "March")
c(TRUE, FALSE)
```

## Variables, missing values and 'factors'

```{r}
age <- c(27, NA, 32, 29)
gender <- factor(
    c("Female", "Male", "Non-binary", NA),
    levels = c("Female", "Male", "Non-binary")
)
```

## Data structures

to coordinate related vectors -- the `data.frame`

```{r}
df <- data.frame(
    age = c(27, NA, 32, 29),
    gender = gender
)
df
```

## Key operations on `data.frame`

```{r}
# Select columns
df$age

# Subset on rows and columns
df[1:3, c("gender", "age")]
```


## Packages

Extend functionality of base *R*. Can be part of the 'base' distribution...

### Where do Packages Come From?

-   \[CRAN\]\[\]: *C*omprehensive *R* *A*rchive *N*etwork. More than 18,000 packages. Some help from \[CRAN Task Views\]\[\] in identifying relevant packages.

-   \[Bioconductor\]\[\]: More than 2100 packages relevant to high-throughput genomic analysis. *Vignettes* are an important part of *Bioconductor* packages.

## Install packages once per *R* installation, using

```{r eval=FALSE}
BiocManager::install(<package-name>)` (CRAN or Bioconductor)
```

## Help & Vignettes

1.  Help pages, e.g., `?lm`

2.  Vignettes, e.g.,

```{r, eval = FALSE}
vignette(package = "ggplot2")
vignette("ggplot2-specs", "ggplot2")
```

3.  Google, StackOverflow, etc...
